### 2016-02-12 mcmc.tex
kill two HOGGs, add one back

% to-do
% -----
% - Kill all issues in this list, below
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Systematically cite chapters of Brooks \etal in each section.
% - Skim MacKay and make sure we cite him in any relevant place.
% - mention Gelman--Rubin statistic
% - mention & cite JAGS and STAN in methods section
% - do we explain how to initialize emcee; do we need to?
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function, by, say, 1-d scans back and forth.
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

% style notes
% -----------
% - \MCMC not MCMC
% - These documents are ``Chapters'', no?  Use \documentname.
% - [LaTeX] newline after every full stop for proper git diffing
% - [LaTeX] eqnarray not equation
% - pdf not PDF
% - is it ``a MCMC sampling'' or ``an MCMC sampling''?
% - \note{} right after word if you are endnoting a phrase,
%   but after the full stop if endnoting the sentence or paragraph.
% - always $\pars$ never $x$; $K$ samples $\pars_k$ and $\pars$ space is $D$-dimensional
% - should we use the word ``walker'' everywhere and do we?

@@

@@

Of course figures are always better than tables, at least in the
printed form of scientific publications.
In a $D$-dimensional parameter space, we recommend plotting all $D$
one-dimensional sample histograms, on a pretty fine binning, and all
$D$-choose-2 two-dimensional histograms (scatter plots) to show the
low-level covariances and non-linearities.
If you are clever, all these plots can be arranged into a lovely and
informative triangle.\note{HOGG cite corner.py and put an example in
  the note!}
These two-dimensional plots do not in any sense contain all the
information in the sampling!  However, they are remarkable for
locating expected and unexpected parameter relationships, and often
invaluable for suggesting re-parameterizations and transformations
that simplify your problem.

@@

As we have discussed above (\sectionname~\ref{sec:MH}),
\MCMC\ methods are proven to be correct ``in the limit''; that is, when
an infinite number of samples have been taken.
When the function $f(\pars)$ has challenging properties, the results
can approach this limit very slowly (or not at all\note{HOGG: discuss
  situations in which \MCMC\ will not work---related to support.}).
For this reason, in the finite duration of a scientific project, most
users of \MCMC\ have one of two (related) problems:
Either \textsl{(1)}~it is taking too long---it is taking too many steps or
executing too many calls of the function $f(\pars)$ to get independent
samples---or \textsl{(2)}~it is not
exploring the full parameter space---there are local optima (or bad
local geometry around those optima) ``trapping'' the algorithm.

@@

>>>>

% to-do
% -----
% - Kill all issues in this list, below
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Systematically cite chapters of Brooks \etal in each section.
% - Skim MacKay and make sure we cite him in any relevant place.
% - Mention & cite JAGS and STAN in methods section
% - Do we explain how to initialize emcee; do we need to?
% - Need to mention case where there is an MC integral INSIDE the likelihood function
% - Need to mention testing the likelihood function, by, say, 1-d scans back and forth.
% - Would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - Send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - Post on arXiv

% low priority
% ------------
% - Make endnote numbers clickable links to endnotes and back.  Relevant ideas here:
%   http://tex.stackexchange.com/questions/8452/making-endnotes-clickable-links-with-hyperref

% style notes
% -----------
% - \MCMC not MCMC
% - These documents are ``Chapters'', no?  Use \documentname.
% - [LaTeX] newline after every full stop for proper git diffing
% - [LaTeX] eqnarray not equation
% - pdf not PDF
% - Is it ``a MCMC sampling'' or ``an MCMC sampling''?
% - \note{} right after word if you are endnoting a phrase,
%   but after the full stop if endnoting the sentence or paragraph.
% - Always $\pars$ never $x$; $K$ samples $\pars_k$ and $\pars$ space is $D$-dimensional
% - Should we use the word ``walker'' everywhere and do we?

@@

HOGG: Introduce and explain the Gelman--Rubin
diagnostic\note{\citet{gelmanrubin}.}, which is also a good idea, and
which tests something slightly different than the autocorrelation
time.

@@

Of course figures are always better than tables, at least in the
printed form of scientific publications.
In a $D$-dimensional parameter space, we recommend plotting all $D$
one-dimensional sample histograms, on a pretty fine binning, and all
$D$-choose-2 two-dimensional histograms (scatter plots) to show the
low-level covariances and non-linearities.
If you are clever, all these plots can be arranged into a lovely and
informative triangle.\note{Our favorite package these days is our own \project{corner.py}.  Find it on the internets.}
These two-dimensional plots do not in any sense contain all the
information in the sampling!  However, they are remarkable for
locating expected and unexpected parameter relationships, and often
invaluable for suggesting re-parameterizations and transformations
that simplify your problem.

@@

As we have discussed above (\sectionname~\ref{sec:MH}),
\MCMC\ methods are proven to be correct ``in the limit''; that is, when
an infinite number of samples have been taken.
When the function $f(\pars)$ has challenging properties, the results
can approach this limit very slowly (or not at all\note{You can get
  essentially infinite autocorrelation times if the function has
  isolated regions of finite density separated by large seas of zero
  or near-zero density.  One disturbing thing about these situations
  is that any empirical measure of autocorrelation time will give a
  finite answer, but the density won't be properly sampled, ever.  In
  principle you can find these with multiple re-starts of the chain
  and (things like) the Gelman--Rubin diagnostic \citep{gelmanrubin}.}).
For this reason, in the finite duration of a scientific project, most
users of \MCMC\ have one of two (related) problems:
Either \textsl{(1)}~it is taking too long---it is taking too many steps or
executing too many calls of the function $f(\pars)$ to get independent
samples---or \textsl{(2)}~it is not
exploring the full parameter space---there are local optima (or bad
local geometry around those optima) ``trapping'' the algorithm.

@@

\bibitem[Gelman \& Rubin(1992)]{gelmanrubin}
  Gelman,~A., \& Rubin,~D.~B., 1992,
  Inference from iterative simulation using multiple sequences,
  \textit{Statist.\ Sci.}\ \textbf{4} 457--472.
